---
title: "Stroke Prediction"
author: "Kartik Ullal"
date: "12/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing all libraries required

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(pscl)
library(ROCR)
library(caret)
library(ROCR)
```

### Reading Dataset


```{r}
dat_stroke = read.csv("healthcare-dataset-stroke-data.csv")
dim(dat_stroke)
```
### Data Cleaning

```{r}
colSums(is.na(dat_stroke))
colSums(dat_stroke == "N/A")
```
#### Checking for duplicate values
```{r}
n_occur <- data.frame(table(dat_stroke$id))
n_occur[n_occur$Freq > 1,]

sum(dat_stroke$gender == "Other")

sum(dat_stroke$smoking_status == "Unknown")
```
#### Filling missing values of smoking status for ages <= 5
```{r}
dat_stroke[dat_stroke$age <= 5, "smoking_status"] <- "never smoked"
```

#### Removing NA values

NA values are insignificant portion of the data, so we will remove them.

```{r}
dat_stroke <- dat_stroke %>%
  filter(
    bmi != "N/A",
    gender != "Other"
  ) %>%
  mutate(
    gender = factor(gender),
    hypertension = factor(hypertension),
    heart_disease = factor(heart_disease),
    ever_married = factor(ever_married),
    work_type = factor(work_type),
    residence_type = factor(Residence_type),
    bmi = as.numeric(bmi),
    smoking_status = factor(smoking_status),
    stroke = factor(stroke),
  ) %>%
  select(
    -Residence_type
  )
```

### Data Visualization

#### Bar plot of parameters with respect to stroke
```{r}
p2 <- dat_stroke %>%
  ggplot(aes(x = hypertension, fill = stroke)) +
  geom_bar()

p3 <- dat_stroke %>%
  ggplot(aes(x = heart_disease, fill = stroke)) +
  geom_bar()

p4 <- dat_stroke %>%
  ggplot(aes(x = ever_married, fill = stroke)) +
  geom_bar()

p7 <- dat_stroke %>%
  ggplot(aes(x = smoking_status, fill = stroke)) +
  geom_bar()

grid.arrange(grobs = list(p2, p3,
                          p4, p7),
             ncol = 2,
             top = "Count of Levels for Each Factor"
)
```

#### Bar plot of proportions of parameters with respect to stroke
```{r}
dat_prop <- dat_stroke %>%
  group_by(hypertension) %>%
  summarise(prop = sum(stroke == "1")/length(hypertension))

p2 <- dat_prop %>%
  ggplot(aes(x = hypertension, y = prop)) +
  geom_col(fill = "#00BFC4")

dat_prop <- dat_stroke %>%
  group_by(heart_disease) %>%
  summarise(prop = sum(stroke == "1")/length(heart_disease))

p3 <- dat_prop %>%
  ggplot(aes(x = heart_disease, y = prop)) +
  geom_col(fill = "#00BFC4")

dat_prop <- dat_stroke %>%
  group_by(ever_married) %>%
  summarise(prop = sum(stroke == "1")/length(ever_married))

p4 <- dat_prop %>%
  ggplot(aes(x = ever_married, y = prop)) +
  geom_col(fill = "#00BFC4")

dat_prop <- dat_stroke %>%
  group_by(smoking_status) %>%
  summarise(prop = sum(stroke == "1")/length(smoking_status))

p7 <- dat_prop %>%
  ggplot(aes(x = smoking_status, y = prop)) +
  geom_col(fill = "#00BFC4")

grid.arrange(grobs = list(p2, p3,
                          p4, p7),
             ncol = 2,
             top = "Proportion of Strokes for Each Factor"
)

```
#### Boxplot of age and parameters with respect to stroke
```{r}
p2 <- dat_stroke %>%
  ggplot(aes(x = hypertension, y = age, color = stroke)) +
  geom_boxplot() +
  theme(legend.position="none")

p3 <- dat_stroke %>%
  ggplot(aes(x = heart_disease, y = age, color = stroke)) +
  geom_boxplot() +
  theme(legend.position="none")

p4 <- dat_stroke %>%
  ggplot(aes(x = ever_married, y = age, color = stroke)) +
  geom_boxplot() +
  theme(legend.position="none")

p7 <- dat_stroke %>%
  ggplot(aes(x = smoking_status, y = age, color = stroke)) +
  geom_boxplot() +
  theme(legend.position="none")

grid.arrange(grobs = list(p2, p3,
                          p4, p7), 
             ncol = 2,
             top = "Stroke and Age Across Factors"
)

```


#### Boxplot of average glucose level and parameters with respect to stroke
```{r}

p2 <- dat_stroke %>%
  ggplot(aes(x = hypertension, y = avg_glucose_level, color = stroke)) +
  geom_boxplot() +
  theme(legend.position="none")

p3 <- dat_stroke %>%
  ggplot(aes(x = heart_disease, y = avg_glucose_level, color = stroke)) +
  geom_boxplot() +
  theme(legend.position="none")

p4 <- dat_stroke %>%
  ggplot(aes(x = ever_married, y = avg_glucose_level, color = stroke)) +
  geom_boxplot() +
  theme(legend.position="none")

p7 <- dat_stroke %>%
  ggplot(aes(x = smoking_status, y = avg_glucose_level, color = stroke)) +
  geom_boxplot() +
  theme(legend.position="none")

grid.arrange(grobs = list(p2, p3,
                          p4, p7), 
             ncol = 2,
             top = "Stroke and Average Glucose Level Across Factors"
)

```

#### Density plot of parameters with stroke

```{r}
p1 <- dat_stroke %>%
  ggplot(aes(x = age, fill = stroke)) +
  geom_density(alpha = 0.5) +
  theme(legend.position="none")

p2 <- dat_stroke %>%
  ggplot(aes(x = avg_glucose_level, fill = stroke)) +
  geom_density(alpha = 0.5) +
  theme(legend.position="none")

p3 <- dat_stroke %>%
  ggplot(aes(x = bmi, fill = stroke)) +
  geom_density(alpha = 0.5) +
  theme(legend.position="none")

p4 <- dat_stroke %>%
  ggplot(aes(x = age, fill = stroke)) +
  geom_histogram() +
  theme(legend.position="none")

p5 <- dat_stroke %>%
  ggplot(aes(x = avg_glucose_level, fill = stroke)) +
  geom_histogram() +
  theme(legend.position="none")

p6 <- dat_stroke %>%
  ggplot(aes(x = bmi, fill = stroke)) +
  geom_histogram() +
  theme(legend.position="none")

grid.arrange(grobs = list(p1, p2, p3,
                          p4, p5, p6), 
             ncol = 3,
             top = "Distribution of Continuous Variables"
)
```

### Logistic Regression Model

#### Creating training and test sets
```{r}
set.seed(2401)
split <- sort(sample(nrow(dat_stroke), nrow(dat_stroke) * 0.7))
dat_stroke_train <- dat_stroke[split,]
dat_stroke_test <- dat_stroke[-split,]
```

#### Fitting the logistic regression model
```{r}
glm_stroke <- glm(
  stroke ~ gender + age + hypertension + heart_disease + ever_married + 
    work_type + residence_type + avg_glucose_level + bmi + smoking_status,
  data = dat_stroke_train,
  family = binomial
)
```

```{r}
summary(glm_stroke)
anova(glm_stroke, test = 'Chisq')
```

#### ROC curve
```{r}
roc_pred <- prediction(
  predictions = glm_stroke$fitted.values,
  labels = dat_stroke_train$stroke
)

roc_perf <- performance(
  roc_pred,
  measure = "tpr",
  x.measure = "fpr"
)

roc_curve <- data.frame(
  Spec = 1 - unlist(roc_perf@x.values),
  Sens = unlist(roc_perf@y.values),
  thresh = unlist(roc_perf@alpha.values)
)

roc_curve$distance <- sqrt((1 - roc_curve$Spec)^2 + (1 - roc_curve$Sens)^2)

opt <- roc_curve %>%
  slice(
    distance %>% which.min()
  )


plot(
  roc_perf,
  main = "Logistic Regression for Strokes"
)

abline(0, 1, col = "grey80")

#Optimal Threshold
abline(v = 1 - opt$Spec, col = "gray80")
abline(h = opt$Sens, col = "gray80")
```

#### Testing model on test data
```{r}
glm_predict <- predict.glm(glm_stroke,
                          newdata =  dat_stroke_test,
                           type = "response",
                           se.fit = FALSE) %>% as_tibble()
glm_predict$pred <- ifelse(glm_predict$value >= opt$thresh, 1, 0)
glm_predict <- glm_predict %>%
    mutate(
    pred = factor(pred)
  )
print(glm_predict)

confusionMatrix(
  data = glm_predict$pred,
  reference = dat_stroke_test$stroke,
  positive = "1"
)
```

#### Confusion Matrix
```{r}
table <- data.frame(confusionMatrix(glm_predict$pred, dat_stroke_test$stroke)$table)

plotTable <- table %>%
  mutate(goodbad = ifelse(table$Prediction == table$Reference, "good", "bad")) %>%
  group_by(Reference) %>%
  mutate(prop = Freq/sum(Freq))

# fill alpha relative to sensitivity/specificity by proportional outcomes within reference groups (see dplyr code above as well as original confusion matrix for comparison)
ggplot(data = plotTable, mapping = aes(x = Reference, y = Prediction, fill = goodbad, alpha = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_manual(values = c(good = "green", bad = "red")) +
  theme_bw() +
  xlim(rev(levels(table$Reference)))
```